---
title: "Hello Observable, Quarto, Shiny"
runtime: shinyrmd
format: html
---

## Observable->R

Here's a minimal demo of observable and R talking to each other inside quarto.

### Observable side

```{observable}
viewof ojsSlider = slider({
  min: 0, 
  max: 40, 
  step: 1, 
  value: 30, 
  title: "An observable slider",
})
```

This following call registers the observable view for shiny, and (if it were working) vice-versa.

```{observable}
//| output: false
shinyInput("ojsSlider")
var1 = shinyOutput("var1")
iris = shinyOutput("shinyIris")
```

From there, we can use `var1` and `iris` directly as observable reactive values.

```{observable}
function transpose(df)
{
  const result = [];
  const l = Object.values(df)[0].length;
  for (let i = 0; i < l; ++i) {
    result.push(
      Object.fromEntries(
        Object.entries(df).map(([key, ar]) => [key, ar[i]])));
  }
  return result;
}
Plot.dot(
  transpose(iris), 
  {
    x: "Sepal.Length", 
    y: "Petal.Length", 
    stroke: "Species"
  }).plot()
```

### R side

This is an output produced by Shiny in the R side.

```{r}
textOutput("ojsSlider", container = p)
```


## Improvements

We could maybe do something like `//| shiny: true` to inspect all
cells automatically, or even some yaml at the top-level to declare
cross-language communication, say:

```
  observable_to_shiny:
    - ojsSlider
    - ...
  shiny_to_observable:
    - var1
    - ...
```


## Joe's demo

The rest of Joe's demo is here.

```{r}
p(class = "observablehq-inputs-button", id = "hello", "Hello")
textOutput("clicks", container = p)
```

```{observable}
viewof myslider = Inputs.range([0, 10], { step: 1 });
var1.slice(0, myslider);
```

## Server code

```{r, context="server"}

output$ojsSlider <- renderText({
  input$ojsSlider
})

observe({
  message(input$var1)
})

output$shinyIris <- reactive(iris)

clicks <- reactiveVal(0L)

r_random <- eventReactive(clicks(), {
  Sys.sleep(0.5)
  runif(10)
})

observeEvent(input$hello, {
  clicks(clicks() + 1L)
})

output$clicks <- renderText({
  clicks()
})

output$insp <- reactive({
  clicks()
  list(runif(3), runif(3))
})

output$var1 <- r_random
```


## Imports, etc

```{observable}
import {slider} from '@jashkenas/inputs'
```
